---
title: "EnsembleLearningFullData"
author: "Xutao Wang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE}
# library(BiocParallel)
# library(caret)
library(dplyr)
library(ggplot2)
# library(ggpubr)
# library(ggridges)
library(glmnet)
library(gridExtra)
library(MultiAssayExperiment)
library(SummarizedExperiment)
library(TBSignatureProfiler)
source("~/Desktop/practice/EnsembleTrainingScript.R")
#source("~/Desktop/curatedTBData/R/TBSignatureProfiling.R")
#source("~/Desktop/curatedTBData/R/subset_curatedTBData.R")
devtools::load_all()
```

# Import Data
```{r, message = FALSE, results = 'hide', eval=FALSE}
objects_list <- curatedTBData(geoAccession = "All")
```

```{r, message = FALSE, results='hide', eval=FALSE}
# Subset studies remove repeated measurement
# Exclude GSE31348 (only PTB) GSE3628 (only PTB) GSE74092(RT-PCR)
# Expect 49 - 3 - 5 + 1 - 6 + 1 = 37
geo_nochange <- c("GSE28623", "GSE29536", "GSE34608", "GSE37250", "GSE39939",
                  "GSE39940", "GSE41055", "GSE50834", "GSE62525", "GSE73408",
                  "GSE83892", "GSE101705", "GSE107731", "GSE107991", "GSE107992",
                  "GSE107104", "GSE112104", "GSEBruno", "GSE25534", "GSE6112")
object_edit <- lapply(geo_nochange, function(x) {
  Mobject <- objects_list[[x]]
  dat <- Mobject[["assay_curated"]]
  if (is.null(dat)) {
    stop("check your assay_curated")
  } 
  SummarizedExperiment(assays = list(assay_curated = dat), 
                       colData = colData(Mobject))
})
names(object_edit) <- geo_nochange

## Combine GSE19491
GSE19491_geo <- c("GSE19435", "GSE19439", "GSE19442", "GSE19444", "GSE22098")

GSE19491_combine <- CombineObjects(objects_list,
                                   list_name = GSE19491_geo,
                                   experiment_name = "assay_curated")

GSE19435_baseline <- objects_list$GSE19435[, objects_list$GSE19435$MeasurementTime
                                           == "0_months"]
GSE19491_edit <- GSE19491_combine[, c(colnames(GSE19435_baseline)[["assay_curated"]],
                                       colnames(objects_list$GSE19439)[["assay_curated"]],
                                       colnames(objects_list$GSE19442)[["assay_curated"]],
                                       colnames(objects_list$GSE19444)[["assay_curated"]],
                                       colnames(objects_list$GSE22098)[["assay_curated"]])]
object_edit$GSE19491_edit <- GSE19491_edit
## Combine GSE42834
GSE42834_geo <- c("GSE42825", "GSE42826", "GSE42827", "GSE42830", "GSE42831", "GSE42832")
GSE42834_combine <- CombineObjects(objects_list,
                                   list_name = GSE42834_geo,
                                   experiment_name = "assay_curated")
GSE42832_sobject_WB <- objects_list$GSE42832[, objects_list$GSE42832$Tissue ==
                                               "Whole Blood"]
GSE42834_edit <- GSE42834_combine[,c(colnames(objects_list$GSE42825)[["assay_curated"]],
                                               colnames(objects_list$GSE42826)[["assay_curated"]],
                                               colnames(objects_list$GSE42827)[["assay_curated"]],
                                               colnames(objects_list$GSE42830)[["assay_curated"]],
                                               colnames(objects_list$GSE42831)[["assay_curated"]],
                                               colnames(GSE42832_sobject_WB)[["assay_curated"]])]
object_edit$GSE42834_edit <- GSE42834_edit

## Subset GSE19443
# Subset samples with cell type: Neutrophils
object_edit$GSE19443_edit <- objects_list$GSE19443[, objects_list$GSE19443$Tissue == "Neutrophils"]

## Subset GSE56153
# Get PTB measurement at Baseline and Controls
GSE56153_baseline <- objects_list$GSE56153[, objects_list$GSE56153$MeasurementTime
                                           %in% c("0 weeks", NA)]
object_edit$GSE56153_edit <- SummarizedExperiment(
  list(counts = GSE56153_baseline[["assay_curated"]]),
  colData = colData(GSE56153_baseline))

## Subset GSE54992
# Only include samples at baseline
GSE54992_baseline <- objects_list$GSE54992[,objects_list$GSE54992$MeasurementTime
                                           == "Baseline"]
object_edit$GSE54992_edit <- SummarizedExperiment(
  list(counts = GSE54992_baseline[["assay_curated"]]),
  colData = colData(GSE54992_baseline))

## GSE62147
# Only include samples prior to treatment
GSE62147_pre_treatment <- objects_list$GSE62147[, objects_list$GSE62147$MeasurementTime
                                                == "recruit"]
object_edit$GSE62147_edit <- SummarizedExperiment(
  list(counts = GSE62147_pre_treatment[["assay_curated"]]),
  colData = colData(GSE62147_pre_treatment))

## GSE69581
# Exclude 10 subclinical samples
GSE69581_PTB_Latent <- objects_list$GSE69581[, objects_list$GSE69581$TBStatus
                                             %in% c("PTB", "LTBI")]
object_edit$GSE69581_edit <- SummarizedExperiment(
  list(counts = GSE69581_PTB_Latent[["assay_curated"]]),
  colData = colData(GSE69581_PTB_Latent))

## GSE79362
# Use the reprocessed RNA-seq, only inlude progressors and non-progressors at baseline
counts.africa.baseline <- objects_list$GSE79362[["assay_reprocess_hg19"]]
# Max 5 filter
MaxFilter <- function(df, max.value = 10){
  df.filtered <- df[which(apply(df,1,max) >= max.value),]
  return(df.filtered)
}
counts.africa.baseline.filtered <- MaxFilter(counts.africa.baseline, 5)
# Normalization
counts.africa.baseline.norm <- TBSignatureProfiler::deseq2_norm_rle(counts.africa.baseline.filtered)
GSE79362_train_full <- SummarizedExperiment(list(counts=counts.africa.baseline.norm),
                                            colData = colData(objects_list$GSE79362))
sample_baseline_GSE79362 <- colData(GSE79362_train_full)[, c("PatientID","MeasurementTime")] %>%
  data.frame() %>%
  dplyr::mutate(sample_name = row.names(colData(GSE79362_train_full))) %>%
  dplyr::arrange(MeasurementTime, PatientID) %>%
  dplyr::group_by(PatientID) %>%
  dplyr::mutate(first = dplyr::first(sample_name))
GSE79362_baseline <- GSE79362_train_full[,unique(sample_baseline_GSE79362$first)]
object_edit$GSE79362_edit <- GSE79362_baseline ## validated with the Brazil data

## Subset GSE81746
# Exclude the pooled PTB (Male)
GSE81746_sub <- objects_list$GSE81746[,objects_list$GSE81746$Gender == "Male"]
object_edit$GSE81746_edit <- SummarizedExperiment(
  list(counts = GSE81746_sub[["assay_curated"]]),
  colData = colData(GSE81746_sub))

## Subset GSE83456
# Exclude patients with EPTB 
GSE83456_sub <- objects_list$GSE83456[,objects_list$GSE83456$TBStatus != "EPTB"]
object_edit$GSE83456_edit <- SummarizedExperiment(
  list(counts = GSE83456_sub[["assay_curated"]]),
  colData = colData(GSE83456_sub))

## GSE84076
# Take BCG vaccinated controls and LTBIs
# Take PTB before treatment results
GSE84076_BCG <- objects_list$GSE84076[, objects_list$GSE84076$BcgVaccinated == "Yes"]
GSE84076_beforeTreat <- objects_list$GSE84076[, objects_list$GSE84076$TreatmentStatus == "Treatment-naive"]
GSE84076_sub1 <- objects_list$GSE84076[,c(colnames(GSE84076_BCG)[["assay_curated"]],
                                          colnames(GSE84076_beforeTreat)[["assay_curated"]])]
object_edit$GSE84076_edit <- SummarizedExperiment(list(counts = GSE84076_sub1[["assay_curated"]]),
                                                   colData = colData(GSE84076_sub1))
## GSE107994
# Use the reprocessed RNA-seq, only include samples at baseline
counts.gse107994.baseline <- objects_list$GSE107994[["assay_reprocess_hg38"]]
# Max 5 filter
counts.gse107994.baseline.filtered <- MaxFilter(counts.gse107994.baseline, 5)
# Normalization
counts.gse107994.baseline.norm <- TBSignatureProfiler::deseq2_norm_rle(counts.gse107994.baseline.filtered)

GSE107994_test_full <- SummarizedExperiment(list(counts=counts.gse107994.baseline.norm),
                                            colData = colData(objects_list$GSE107994))
# index <- which(is.na(GSE107994_test_full$PatientID))
sample_baseline_GSE107994 <- colData(GSE107994_test_full)[, c("PatientID","Progression")] %>%
  data.frame() %>%
  dplyr::mutate(sample_name = row.names(colData(GSE107994_test_full))) %>%
  dplyr::group_by(PatientID) %>%
  dplyr::mutate(first = dplyr::first(sample_name))

# Patient_087 does not have baseline measurement
GSE107994_baseline <- GSE107994_test_full[, unique(sample_baseline_GSE107994$first)]

object_edit$GSE107994_edit <- GSE107994_baseline

## GSE94438
# Use reprocessed RNA-seq, exclude samples with TBStatus NA
counts.gse94438.baseline <- objects_list$GSE94438[["assay_reprocess_hg19"]]
# Max 5 filter
counts.gse94438.baseline.filtered <- MaxFilter(counts.gse94438.baseline, 5)
# Normalization
counts.gse94438.baseline.norm <- TBSignatureProfiler::deseq2_norm_rle(counts.gse94438.baseline.filtered)

GSE94438_test_full <- SummarizedExperiment(list(counts=counts.gse94438.baseline.norm),
                                           colData = colData(objects_list$GSE94438))
attributes(row.names(GSE94438_test_full)) <- NULL
GSE94438_test_full_NoNA <- GSE94438_test_full[,GSE94438_test_full$Progression %in% c("Positive","Negative")]
object_edit$GSE94438_edit <- GSE94438_test_full_NoNA

## GSE89403
# Only include samples collected at baseline, or when it is measured at first time
# Remove samples with TBStatus NA
sample_baseline_GS89403 <- colData(objects_list$GSE89403)[, c("PatientID","MeasurementTime")] %>%
  data.frame() %>%
  dplyr::mutate(sample_name = row.names(colData(objects_list$GSE89403))) %>%
  dplyr::arrange(MeasurementTime, PatientID) %>%
  dplyr::group_by(PatientID) %>%
  dplyr::mutate(first = dplyr::first(sample_name))
GS89403_baseline <- objects_list$GSE89403[,unique(sample_baseline_GS89403$first)]
GS89403_baseline_noNA <- GS89403_baseline[, !is.na(GS89403_baseline$TBStatus)]
object_edit$GSE89403_edit <- SummarizedExperiment(
  list(counts = GS89403_baseline_noNA[["assay_curated"]]),
  colData = colData(GS89403_baseline_noNA))

## GSE107993
# Only include samples collected at baseline, or when it is measured at first time
sample_baseline_GSE107993 <- colData(objects_list$GSE107993)[, c("PatientID","MeasurementTime")] %>%
  data.frame() %>%
  dplyr::mutate(sample_name = row.names(colData(objects_list$GSE107993))) %>%
  dplyr::arrange(MeasurementTime, PatientID) %>%
  dplyr::group_by(PatientID) %>%
  dplyr::mutate(first = dplyr::first(sample_name))
GSE107993_baseline <- objects_list$GSE107993[, unique(sample_baseline_GSE107993$first)]

object_edit$GSE107993_edit <- SummarizedExperiment(
  list(counts = GSE107993_baseline[["assay_curated"]]),
  colData = colData(GSE107993_baseline)[colnames(GSE107993_baseline[["assay_curated"]]),])

## GSETornheim
# Subset samples collected at baseline, or when it is measured at the first time
# Remove patients with EPTB
sample_baseline_GSETornheim <- colData(objects_list$GSETornheim)[, c("PatientID", "MeasurementTime")] %>%
  data.frame() %>%
  dplyr::mutate(sample_name = row.names(colData(objects_list$GSETornheim))) %>%
  dplyr::arrange(MeasurementTime) %>% 
  dplyr::group_by(PatientID) %>%
  dplyr::mutate(first = dplyr::first(sample_name))
sample_baseline_GSEToenheim <- objects_list$GSETornheim[, unique(sample_baseline_GSETornheim$first)]
sample_baseline_GSEToenheimSub <- sample_baseline_GSEToenheim[, sample_baseline_GSEToenheim$TBStatus %in% c("Control", "PTB")]
object_edit$GSETornheim_edit <- SummarizedExperiment(
  list(counts = sample_baseline_GSEToenheimSub[["assay_curated"]]),
  colData = colData(sample_baseline_GSEToenheimSub))

## Subset GSE40553
# Subset samples collected at baseline, or when it is measured at the first time
sample_baseline_GSE40553 <- colData(objects_list$GSE40553)[, c("PatientID","MeasurementTime")] %>%
  data.frame() %>%
  dplyr::mutate(sample_name = row.names(colData(objects_list$GSE40553))) %>%
  dplyr::arrange(MeasurementTime, PatientID) %>%
  dplyr::group_by(PatientID) %>%
  dplyr::mutate(first = dplyr::first(sample_name))
GSE40553_baseline <- objects_list$GSE40553[, unique(sample_baseline_GSE40553$first)]
object_edit$GSE40553_edit <- SummarizedExperiment(
  list(counts = GSE40553_baseline[["assay_curated"]]),
  colData = colData(GSE40553_baseline))

```
```{r}
# saveRDS(object_edit, "~/Desktop/object_edit.RDS")
object_edit <- readRDS("~/Desktop/object_edit.RDS")
# SignatureInfoTraining <- load("~/Desktop/curatedTBData/data/SignatureInfoTraining.rda")
```

# PTB vs. Control {.tabset}
```{r, message=FALSE, results='hide', warning=FALSE}
# 25 in total
multi_set_PTB_Control <- lapply(object_edit, function(x)
  subset_curatedTBData(x, annotationColName = "TBStatus", 
                       annotationCondition = c("Control","PTB"), 
                       useAssay = assayNames(x),
                       experiment_name = "assay_curated")) %>% plyr::compact()
ssgsea_PTB_Control <- lapply(multi_set_PTB_Control,
                    function(x) TBSignatureProfiler::runTBsigProfiler(
                                  input = x,
                                  useAssay = assayNames(x),
                                  signatures = TBsignatures, 
                                  algorithm = "ssGSEA",
                                  combineSigAndAlgorithm = TRUE))
ssgsea_PTB_Control_combine <- combine_auc(ssgsea_PTB_Control, 
                                annotationColName = "TBStatus", 
                                signatureColNames = names(TBsignatures), 
                                num.boot = NULL, percent = 0.95)
ssgsea_PTB_Control_combine$GSE <- factor(ssgsea_PTB_Control_combine$GSE,
                                         levels = unique(ssgsea_PTB_Control_combine$GSE))
plage_PTB_Control <- lapply(multi_set_PTB_Control,
                    function(x) TBSignatureProfiler::runTBsigProfiler(
                                  input = x,
                                  useAssay = assayNames(x),
                                  signatures = TBsignatures, 
                                  algorithm = "PLAGE",
                                  combineSigAndAlgorithm = TRUE))
plage_PTB_Control_combine <- combine_auc(plage_PTB_Control, 
                                annotationColName = "TBStatus", 
                                signatureColNames = names(TBsignatures), 
                                num.boot = NULL, percent = 0.95)
plage_PTB_Control_combine$GSE <- factor(plage_PTB_Control_combine$GSE,
                                        levels = unique(ssgsea_PTB_Control_combine$GSE))
```

```{r}
mySignatures <- TBsignatures[-which(names(TBsignatures) %in% 
                                      c("Gliddon_OD_3", "Hoang_OD_3",
                                        "LauxdaCosta_OD_3",
                                        "Lee_4", "Rajan_HIV_5", "Roe_3",
                                        "Suliman_RISK_2"))]
ensembleSignaturesList <- lapply(1:40, function(x) {
  set.seed(x)
  index <- sample(length(names(mySignatures)), 10)
  names(mySignatures)[index]
})
names(ensembleSignaturesList) <- paste0("Set", seq_len(length(ensembleSignaturesList)))
# testWithScore_PTB_Control_ssGSEA <- testWithScore(ensembleSignaturesList, 
#                                                   theObjectList = ssgsea_PTB_Control,
#                                                   times = 200, method = "ssGSEA")
# testWithScore_PTB_Control_PLAGE <- testWithScore(ensembleSignaturesList, 
#                                                   theObjectList = plage_PTB_Control,
#                                                   times = 200, method = "PLAGE")
testWithScore_PTB_Control_ssGSEA <- readRDS("~/Desktop/practice/testWithScore_PTB_Control_ssGSEA.RDS")
testWithScore_PTB_Control_PLAGE <- readRDS("~/Desktop/practice/testWithScore_PTB_Control_PLAGE.RDS")
```

## ssGSEA {.tabset}

### Heatmap
```{r, fig.width=13, fig.height=18, fig.wide=TRUE}
# Import signature and data information
heatmap_auc(ssgsea_PTB_Control_combine, SignatureInfoTraining, 
            names(TBsignatures), facet = TRUE, clustering = FALSE) + 
  ggtitle("Heatmap of AUC for PTB vs. Control using ssGSEA")
```

### Coefficient
```{r, fig.width=10, fig.height=10}
Coef_PTB_Control_ssGSEA <- getCoef(testWithScore_PTB_Control_ssGSEA, ensembleSignaturesList)
Coef_PTB_Control_ssGSEA_Df <- do.call(rbind, Coef_PTB_Control_ssGSEA)
Coef_PTB_Control_ssGSEA_Df$Coefficient <- as.numeric(Coef_PTB_Control_ssGSEA_Df$Coefficient)
# Making plot
ggplot(Coef_PTB_Control_ssGSEA_Df, aes(x = Coefficient, y = Signature)) + geom_boxplot() +
  theme_bw() + ggtitle("Coefficient Distribution using ssGSEA (PTB vs. Control)") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

### Analysis of Set
```{r}
AUC_PTB_Control <- readRDS("~/Desktop/practice/AUC_PTB_Control.RDS")
outperform_Control_ssGSEA <- compareAUC(AUC_PTB_Control$AUC_final, 
                                 ensembleSignaturesList, "ssGSEA")
DT::datatable(outperform_Control_ssGSEA)
# Find Sets with combine > individual and display its coefficient
```

```{r, results='hide'}
for (i in outperform_Control_ssGSEA$SetName) {
  cat("####", i, "\n")
  
  coefTab <- Coef_PTB_Control_ssGSEA[[i]] %>%  
    dplyr::mutate(Coef = as.numeric(Coefficient)) %>% 
    dplyr::group_by(Signature) %>% 
    dplyr::summarise(mean = round(mean(Coef), 4))
  print(DT::datatable(coefTab))
  
  cat("\n\n")
}
```

## PLAGE {.tabset}

### Heatmap
```{r, fig.width=13, fig.height=18, fig.wide=TRUE, warning=FALSE}
heatmap_auc(plage_PTB_Control_combine, SignatureInfoTraining, 
            names(TBsignatures), facet = TRUE, clustering = FALSE) + 
  ggtitle("Heatmap of AUC for PTB vs. Control using PLAGE")
```

### Coefficient
```{r, fig.width=10, fig.height=10}
Coef_PTB_Control_PLAGE <- getCoef(testWithScore_PTB_Control_PLAGE, ensembleSignaturesList)
Coef_PTB_Control_PLAGE_Df <- do.call(rbind, Coef_PTB_Control_PLAGE)
Coef_PTB_Control_PLAGE_Df$Coefficient <- as.numeric(Coef_PTB_Control_PLAGE_Df$Coefficient)
# Making plot
ggplot(Coef_PTB_Control_PLAGE_Df, aes(x = Coefficient, y = Signature)) + geom_boxplot() +
  theme_bw() + ggtitle("Coefficient Distribution using PLAGE (PTB vs. Control)") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

### Analysis of each combination
```{r}
# AUC_PTB_Control <- readRDS("~/Desktop/practice/AUC_PTB_Control.RDS")
outperform_Control_PLAGE <- compareAUC(AUC_PTB_Control$AUC_final, 
                                       ensembleSignaturesList, "PLAGE")
DT::datatable(outperform_Control_PLAGE)
# Find Sets with combine > individual and display its coefficient
```

```{r, results='hide'}
for (i in outperform_Control_PLAGE$SetName) {
  cat("####", i, "\n")
  
  coefTab <- Coef_PTB_Control_PLAGE[[i]] %>%  
    dplyr::mutate(Coef = as.numeric(Coefficient)) %>% 
    dplyr::group_by(Signature) %>% 
    dplyr::summarise(mean = round(mean(Coef), 4))
  print(DT::datatable(coefTab))
  
  cat("\n\n")
}
```

## Violin Plot
```{r, fig.width=14, fig.height=10}
# AUC_PTB_Control <- make_violin_plot(ensembleSignaturesList, 
#                                     testWithScore_PTB_Control_ssGSEA,
#                                     testWithScore_PTB_Control_PLAGE)
# saveRDS(AUC_PTB_Control,"~/Desktop/practice/AUC_PTB_Control.RDS")
# AUC_PTB_Control <- readRDS("~/Desktop/practice/AUC_PTB_Control.RDS")
p_control <- show_voilin_plot(AUC_PTB_Control, y = 0.75)
p_control + theme(axis.text.x = element_text(size = 12),
                  axis.text.y = element_text(size = 12))
```

# PTB vs. LTBI {.tabset}
```{r, message=FALSE, results='hide', warning=FALSE}
# 25 in total
multi_set_PTB_LTBI <- lapply(object_edit, function(x)
  subset_curatedTBData(x, annotationColName = "TBStatus", 
                       annotationCondition = c("PTB","LTBI"), 
                       useAssay = assayNames(x),
                       experiment_name = "assay_curated")) %>% plyr::compact()
ssgsea_PTB_LTBI <- lapply(multi_set_PTB_LTBI,
                    function(x) TBSignatureProfiler::runTBsigProfiler(
                                  input = x,
                                  useAssay = assayNames(x),
                                  signatures = TBsignatures, 
                                  algorithm = "ssGSEA",
                                  combineSigAndAlgorithm = TRUE))
ssgsea_PTB_LTBI_combine <- combine_auc(ssgsea_PTB_LTBI, 
                                annotationColName = "TBStatus", 
                                signatureColNames = names(TBsignatures), 
                                num.boot = NULL, percent = 0.95)
ssgsea_PTB_LTBI_combine$GSE <- factor(ssgsea_PTB_LTBI_combine$GSE,
                                         levels = unique(ssgsea_PTB_LTBI_combine$GSE))
plage_PTB_LTBI <- lapply(multi_set_PTB_LTBI,
                    function(x) TBSignatureProfiler::runTBsigProfiler(
                                  input = x,
                                  useAssay = assayNames(x),
                                  signatures = TBsignatures, 
                                  algorithm = "PLAGE",
                                  combineSigAndAlgorithm = TRUE))
plage_PTB_LTBI_combine <- combine_auc(plage_PTB_LTBI, 
                                annotationColName = "TBStatus", 
                                signatureColNames = names(TBsignatures), 
                                num.boot = NULL, percent = 0.95)
plage_PTB_LTBI_combine$GSE <- factor(plage_PTB_LTBI_combine$GSE,
                                        levels = unique(ssgsea_PTB_LTBI_combine$GSE))
```

```{r}
testWithScore_PTB_LTBI_ssGSEA <- readRDS("~/Desktop/practice/testWithScore_PTB_LTBI_ssGSEA.RDS")
testWithScore_PTB_LTBI_PLAGE <- readRDS("~/Desktop/practice/testWithScore_PTB_LTBI_PLAGE.RDS")

# AUC_ssGSEA_sig_PTB_LTBI <- testWithScoreDataProcess(testWithScore_PTB_LTBI_ssGSEA, 
#                                                    "Combine_ssGSEA", "ssGSEA",
#                                                     combine = FALSE)
# AUC_plage_sig_PTB_LTBI <- testWithScoreDataProcess(testWithScore_PTB_LTBI_PLAGE, 
#                                                    "Combine_PLAGE", "PLAGE", 
#                                                     combine = FALSE)
# AUC_sig_PTB_LTBI <- rbind(do.call(rbind, AUC_ssGSEA_sig_PTB_LTBI),
#                           do.call(rbind, AUC_plage_sig_PTB_LTBI))
```

## ssGSEA {.tabset}

### Heatmap
```{r, fig.width=13, fig.height=18, fig.wide=TRUE}
# Import signature and data information
heatmap_auc(ssgsea_PTB_LTBI_combine, SignatureInfoTraining, 
            names(TBsignatures), facet = TRUE, clustering = FALSE) + 
  ggtitle("Heatmap of AUC for PTB vs. LTBI using ssGSEA")
```

### Coefficient

```{r, fig.width=10, fig.height=10}
Coef_PTB_LTBI_ssGSEA <- getCoef(testWithScore_PTB_LTBI_ssGSEA, ensembleSignaturesList)
Coef_PTB_LTBI_ssGSEA_Df <- do.call(rbind, Coef_PTB_LTBI_ssGSEA)
Coef_PTB_LTBI_ssGSEA_Df$Coefficient <- as.numeric(Coef_PTB_LTBI_ssGSEA_Df$Coefficient)
# Making plot
ggplot(Coef_PTB_LTBI_ssGSEA_Df, aes(x = Coefficient, y = Signature)) + geom_boxplot() +
  theme_bw() + ggtitle("Coefficient Distribution using ssGSEA (PTB vs. LTBI)") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

### Analysis of Set
```{r}
AUC_PTB_LTBI <- readRDS("~/Desktop/practice/AUC_PTB_LTBI.RDS")
outperform_LTBI_ssGSEA <- compareAUC(AUC_PTB_LTBI$AUC_final, 
                                 ensembleSignaturesList, "ssGSEA")
DT::datatable(outperform_LTBI_ssGSEA)
# Find Sets with combine > individual and display its coefficient
```

```{r, results='hide'}
for (i in outperform_LTBI_ssGSEA$SetName) {
  cat("####", i, "\n")
  
  coefTab <- Coef_PTB_LTBI_ssGSEA[[i]] %>%  
    dplyr::mutate(Coef = as.numeric(Coefficient)) %>% 
    dplyr::group_by(Signature) %>% 
    dplyr::summarise(mean = round(mean(Coef), 4))
  print(DT::datatable(coefTab))
  
  cat("\n\n")
}
```

## PLAGE {.tabset}

### Heatmap
```{r, fig.width=13, fig.height=18, fig.wide=TRUE, warning=FALSE}
heatmap_auc(plage_PTB_LTBI_combine, SignatureInfoTraining, 
            names(TBsignatures), facet = TRUE, clustering = FALSE) + 
  ggtitle("Heatmap of AUC for PTB vs. LTBI using PLAGE")
```

### Coefficient
```{r, fig.width=10, fig.height=10}
Coef_PTB_LTBI_PLAGE <- getCoef(testWithScore_PTB_LTBI_PLAGE, ensembleSignaturesList)
Coef_PTB_LTBI_PLAGE_Df <- do.call(rbind, Coef_PTB_LTBI_PLAGE)
Coef_PTB_LTBI_PLAGE_Df$Coefficient <- as.numeric(Coef_PTB_LTBI_PLAGE_Df$Coefficient)
# Making plot
ggplot(Coef_PTB_LTBI_PLAGE_Df, aes(x = Coefficient, y = Signature)) + geom_boxplot() +
  theme_bw() + ggtitle("Coefficient Distribution using PLAGE (PTB vs. LTBI)") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

### Analysis of Set
```{r}
# AUC_PTB_LTBI <- readRDS("~/Desktop/practice/AUC_PTB_LTBI.RDS")
outperform_LTBI_PLAGE <- compareAUC(AUC_PTB_LTBI$AUC_final, 
                                 ensembleSignaturesList, "PLAGE")
DT::datatable(outperform_LTBI_PLAGE)
# Find Sets with combine > individual and display its coefficient
```

```{r, results='hide'}
for (i in outperform_LTBI_PLAGE$SetName) {
  cat("####", i, "\n")
  
  coefTab <- Coef_PTB_LTBI_PLAGE[[i]] %>%  
    dplyr::mutate(Coef = as.numeric(Coefficient)) %>% 
    dplyr::group_by(Signature) %>% 
    dplyr::summarise(mean = round(mean(Coef), 4))
  print(DT::datatable(coefTab))
  
  cat("\n\n")
}
```

## Violin Plot
```{r, fig.width=14, fig.height=10}
# AUC_PTB_LTBI <- make_violin_plot(ensembleSignaturesList, 
#                                     testWithScore_PTB_LTBI_ssGSEA,
#                                     testWithScore_PTB_LTBI_PLAGE)
# saveRDS(AUC_PTB_LTBI, "~/Desktop/practice/AUC_PTB_LTBI.RDS")
AUC_PTB_LTBI <- readRDS("~/Desktop/practice/AUC_PTB_LTBI.RDS")
p_LTBI <- show_voilin_plot(AUC_PTB_LTBI, y = 0.8)
p_LTBI + theme(axis.text.x = element_text(size = 12),
               axis.text.y = element_text(size = 12))
```

# PTB vs. OD {.tabset}
```{r, message=FALSE, results='hide', warning=FALSE}
# 25 in total
multi_set_PTB_OD <- lapply(object_edit, function(x)
  subset_curatedTBData(x, annotationColName = "TBStatus", 
                       annotationCondition = c("PTB","OD"), 
                       useAssay = assayNames(x),
                       experiment_name = "assay_curated")) %>% plyr::compact()
ssgsea_PTB_OD <- lapply(multi_set_PTB_OD,
                    function(x) TBSignatureProfiler::runTBsigProfiler(
                                  input = x,
                                  useAssay = assayNames(x),
                                  signatures = TBsignatures, 
                                  algorithm = "ssGSEA",
                                  combineSigAndAlgorithm = TRUE))
ssgsea_PTB_OD_combine <- combine_auc(ssgsea_PTB_OD, 
                                annotationColName = "TBStatus", 
                                signatureColNames = names(TBsignatures), 
                                num.boot = NULL, percent = 0.95)
ssgsea_PTB_OD_combine$GSE <- factor(ssgsea_PTB_OD_combine$GSE,
                                         levels = unique(ssgsea_PTB_OD_combine$GSE))
plage_PTB_OD <- lapply(multi_set_PTB_OD,
                    function(x) TBSignatureProfiler::runTBsigProfiler(
                                  input = x,
                                  useAssay = assayNames(x),
                                  signatures = TBsignatures, 
                                  algorithm = "PLAGE",
                                  combineSigAndAlgorithm = TRUE))
plage_PTB_OD_combine <- combine_auc(plage_PTB_OD, 
                                annotationColName = "TBStatus", 
                                signatureColNames = names(TBsignatures), 
                                num.boot = NULL, percent = 0.95)
plage_PTB_OD_combine$GSE <- factor(plage_PTB_OD_combine$GSE,
                                        levels = unique(ssgsea_PTB_OD_combine$GSE))
```

```{r}
testWithScore_PTB_OD_ssGSEA <- readRDS("~/Desktop/practice/testWithScore_PTB_OD_ssGSEA.RDS")
testWithScore_PTB_OD_PLAGE <- readRDS("~/Desktop/practice/testWithScore_PTB_OD_PLAGE.RDS")

# AUC_ssGSEA_sig_PTB_OD <- testWithScoreDataProcess(testWithScore_PTB_OD_ssGSEA, 
#                                                        "Combine_ssGSEA", "ssGSEA",
#                                                         combine = FALSE)
# AUC_plage_sig_PTB_OD <- testWithScoreDataProcess(testWithScore_PTB_OD_PLAGE, 
#                                                       "Combine_PLAGE", "PLAGE", 
#                                                        combine = FALSE)
# AUC_sig_PTB_OD <- rbind(do.call(rbind, AUC_ssGSEA_sig_PTB_OD),
#                         do.call(rbind, AUC_plage_sig_PTB_OD))
```

## ssGSEA {.tabset}
### Heatmap
```{r, fig.width=13, fig.height=18, fig.wide=TRUE}
# Import signature and data information
heatmap_auc(ssgsea_PTB_OD_combine, SignatureInfoTraining, 
            names(TBsignatures), facet = TRUE, clustering = FALSE) + 
  ggtitle("Heatmap of AUC for PTB vs. LTBI using ssGSEA")
```

### Coefficient
```{r, fig.width=10, fig.height=10}
Coef_PTB_OD_ssGSEA <- getCoef(testWithScore_PTB_OD_ssGSEA, ensembleSignaturesList)
Coef_PTB_OD_ssGSEA_Df <- do.call(rbind, Coef_PTB_OD_ssGSEA)
Coef_PTB_OD_ssGSEA_Df$Coefficient <- as.numeric(Coef_PTB_OD_ssGSEA_Df$Coefficient)
# Making plot
ggplot(Coef_PTB_OD_ssGSEA_Df, aes(x = Coefficient, y = Signature)) + geom_boxplot() +
  theme_bw() + ggtitle("Coefficient Distribution using ssGSEA (PTB vs. OD)") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

### Analysis of Set
```{r}
AUC_PTB_OD <- readRDS("~/Desktop/practice/AUC_PTB_OD.RDS")
outperform_OD_ssGSEA <- compareAUC(AUC_PTB_OD$AUC_final, 
                                 ensembleSignaturesList, "ssGSEA")
DT::datatable(outperform_OD_ssGSEA)
# Find Sets with combine > individual and display its coefficient
```

```{r, results='hide'}
for (i in outperform_OD_ssGSEA$SetName) {
  cat("####", i, "\n")
  
  coefTab <- Coef_PTB_OD_ssGSEA[[i]] %>%  
    dplyr::mutate(Coef = as.numeric(Coefficient)) %>% 
    dplyr::group_by(Signature) %>% 
    dplyr::summarise(mean = round(mean(Coef), 4))
  print(DT::datatable(coefTab))
  
  cat("\n\n")
}
```

## PLAGE {.tabset}
### Heatmap
```{r, fig.width=10, fig.height=15, fig.wide=TRUE, warning=FALSE}
heatmap_auc(plage_PTB_OD_combine, SignatureInfoTraining, 
            names(TBsignatures), facet = TRUE, clustering = FALSE) + 
  ggtitle("Heatmap of AUC for PTB vs. OD using PLAGE")
```

### Coefficient
```{r, fig.width=10, fig.height=10}
Coef_PTB_OD_PLAGE <- getCoef(testWithScore_PTB_OD_PLAGE, ensembleSignaturesList)
Coef_PTB_OD_PLAGE_Df <- do.call(rbind, Coef_PTB_OD_PLAGE)
Coef_PTB_OD_PLAGE_Df$Coefficient <- as.numeric(Coef_PTB_OD_PLAGE_Df$Coefficient)
# Making plot
ggplot(Coef_PTB_OD_PLAGE_Df, aes(x = Coefficient, y = Signature)) + geom_boxplot() +
  theme_bw() + ggtitle("Coefficient Distribution using PLAGE (PTB vs. OD)") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

### Analysis of Set
```{r}
# AUC_PTB_OD <- readRDS("~/Desktop/practice/AUC_PTB_OD.RDS")
outperform_OD_PLAGE <- compareAUC(AUC_PTB_OD$AUC_final, 
                                 ensembleSignaturesList, "PLAGE")
DT::datatable(outperform_OD_PLAGE)
# Find Sets with combine > individual and display its coefficient
```

```{r, results='hide'}
for (i in outperform_OD_ssGSEA$SetName) {
  cat("####", i, "\n")
  
  coefTab <- Coef_PTB_OD_PLAGE[[i]] %>%  
    dplyr::mutate(Coef = as.numeric(Coefficient)) %>% 
    dplyr::group_by(Signature) %>% 
    dplyr::summarise(mean = round(mean(Coef), 4))
  print(DT::datatable(coefTab))
  
  cat("\n\n")
}
```

## Violin Plot
```{r, fig.width=14, fig.height=10}
# AUC_PTB_OD <- make_violin_plot(ensembleSignaturesList, 
#                                testWithScore_PTB_OD_ssGSEA,
#                                testWithScore_PTB_OD_PLAGE)
# saveRDS(AUC_PTB_OD, "~/Desktop/practice/AUC_PTB_OD.RDS")
AUC_PTB_OD <- readRDS("~/Desktop/practice/AUC_PTB_OD.RDS")
p_OD <- show_voilin_plot(AUC_PTB_OD, y = 0.68)
p_OD + theme(axis.text.x = element_text(size = 12),
             axis.text.y = element_text(size = 12))
```

