---
title: "vignette-TBSignature"
author: "Xutao Wang"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{vignette-TBSignature}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache.lazy = FALSE)
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
if (! "SummarizedExperiment" %in% installed.packages()) BiocManager::install("SummarizedExperiment")
if (! "MultiAssayExperiment" %in% installed.packages()) BiocManager::install("MultiAssayExperiment")
if (! "GEOquery" %in% installed.packages()) BiocManager::install("GEOquery")
if (! "dplyr" %in% installed.packages()) install.packages("dplyr")
if (! "plyr" %in% installed.packages()) install.packages("plyr")
if (! "TBSignatureProfiler" %in% installed.packages()) devtools::install_github("aubreyodom/TBSignatureProfiler")
```

```{r libraries, message = FALSE}
library(GEOquery)
library(SummarizedExperiment)

#library(limma)
library(dplyr)
library(plyr)
devtools::load_all(".")
library(curatedTBData)
library(TBSignatureProfiler)
library(ggplot2)
library(gridExtra)
library(ggridges)

library(magick)
# library(illuminaHumanv4.db) # Illumina HumanHT-12 V4.0 expression beadchip
```

# Load lists of summarized experiment objects

```{r load list of sobject, message=FALSE}
# Load data
list_files <- data(package="curatedTBData")[[3]][,3]
data(list=list_files) # Takes some time...
strSobject <- ls(pattern="*_sobject$")

# Get a list of Summarized Experiment Object
sobjects_list <- lapply(strSobject, function(x) get(x))
names(sobjects_list) <- gsub("_.*","",strSobject)

# Create Multi-Assay Experiment object Take SOME TIME
multi_data_all <- lapply(sobjects_list, function(x) Create_MultiAssay_object(x))

```

# Visualization with TBsignatures: Comparison across TB Status {.tabset}

## Positive TB (PTB) vs. Latent TB (Latent){.tabset}

```{r calculate ssgsea for PTB vs Latent, message = FALSE, results = "hide"}
multi_set_PTB_Latent <- lapply(multi_data_all, function(x) get_sobject_TBSig(x,"PTB","Latent")) %>% plyr::compact()

ssgsea_PTB_Latent <- lapply(multi_set_PTB_Latent,
                    function(x) TBSignatureProfiler::runTBsigProfiler(input = x,
                                  useAssay = "counts",
                                  signatures = TBsignatures,
                                  algorithm = "ssGSEA",
                                  combineSigAndAlgorithm = TRUE,
                                  parallel.sz = 10))

# No Sloot_HIV_2, Suliman_RISK_4 in some datasets
TBsig_PTB_Latent <- names(TBsignatures)[-grep("Sloot_HIV_2|Suliman_RISK_4",names(TBsignatures))]

index_Sloot <- which(!sapply(ssgsea_PTB_Latent, function(x) "Sloot_HIV_2" %in% names(colData(x))))
index_Suliman <- which(!sapply(ssgsea_PTB_Latent, function(x) "Suliman_RISK_4" %in% names(colData(x))))
# Define new function for box plot display
get_boxplot <- function(result_list,sig_name){
  
  p_boxplot <- lapply(1:length(result_list), function(x)
    TBSignatureProfiler::signatureBoxplot(inputData = result_list[[x]],
                                      name = names(result_list)[x],
                                      signatureColNames = sig_name,
                                      annotationColName = "Disease", rotateLabels = FALSE,fill_colors = c("#00AFBB", "#E7B800")))

    library(gridExtra)
    library(ggplot2)
    p_combine <- do.call("grid.arrange", c(p_boxplot, ncol=floor(sqrt(length(p_boxplot)))))
    return(p_combine)

}
#p = get_boxplot(ssgsea_PTB_Latent, sig_name = "Sweeney_OD_3")
#ggsave("~/Desktop/PresentationFigure/Sweeney_Latent.pdf",p, width = 24, height = 24, 
#       units = "in", # other options are "in", "cm", "mm" 
#       dpi = 200)
p = get_boxplot(ssgsea_PTB_Latent, sig_name = "Anderson_42")
ggsave("~/Desktop/PresentationFigure/Anderson_Latent.pdf",p, width = 24, height = 24, 
       units = "in", # other options are "in", "cm", "mm" 
       dpi = 200)
```

### Boxplot of all scores signatures{.tabset}

```{r Boxplot for each signature PTB vs. Latent,results = "asis", message = FALSE, fig.height = 20, fig.width=28}
for (i in TBsig_PTB_Latent){

  cat("####", i, "\n")
  
  get_boxplot(ssgsea_PTB_Latent, sig_name = i)
  #print(signatureBoxplot(ssgsea_result, name = i, signatureColNames = i,
  #               annotationColName = c("Disease")))
  
  cat("\n\n")
}
```

#### Sloot_HIV_2

```{r boxplot for Sloot_HIV_2 PTB vs. Latent, results="hide",fig.height = 16, fig.width=24}
get_boxplot(ssgsea_PTB_Latent[-index_Sloot], sig_name = "Sloot_HIV_2")

```

#### Suliman_RISK_4

```{r boxplot for Suliman_RISK_4 PTB vs. Latent, results="hide",fig.height = 16, fig.width=24}
get_boxplot(ssgsea_PTB_Latent[-index_Suliman], sig_name = "Suliman_RISK_4")

```

## Positive TB (PTB) vs. Health Control (Control){.tabset}

```{r calculate ssgsea for PTB vs Control, message = FALSE, results = "hide"}
multi_set_PTB_Control <- lapply(multi_data_all, function(x) get_sobject_TBSig(x,"PTB","Control")) %>% plyr::compact()

ssgsea_PTB_Control <- lapply(multi_set_PTB_Control,
                                 function(x) runTBsigProfiler(input = x,
                                                              useAssay = "counts",
                                                              signatures = TBsignatures,
                                                              algorithm = "ssGSEA",
                                                              combineSigAndAlgorithm = TRUE,
                                                              parallel.sz = 10))

TBsig_PTB_Control <- names(TBsignatures)[-which(names(TBsignatures)%in% c("Lee_4","Rajan_HIV_5","Sloot_HIV_2","Thompson_FAIL_13","Suliman_RISK_4"))]

#  "Lee_4","Rajan_HIV_5","Sloot_HIV_2","Thompson_FAIL_13" are in the same gene set
index_Lee <- which(!sapply(ssgsea_PTB_Control, function(x) "Lee_4" %in% names(colData(x))))
index_Suliman <- which(!sapply(ssgsea_PTB_Control, function(x) "Suliman_RISK_4" %in% names(colData(x))))


p = get_boxplot(ssgsea_PTB_Control, sig_name = "Sweeney_OD_3")
ggsave("~/Desktop/PresentationFigure/Sweeney_Control.pdf",p, width = 24, height = 24, 
       units = "in", # other options are "in", "cm", "mm" 
       dpi = 200)

p = get_boxplot(ssgsea_PTB_Control, sig_name = "Anderson_42")
ggsave("~/Desktop/PresentationFigure/Anderson_Control.pdf",p, width = 24, height = 24, 
       units = "in", # other options are "in", "cm", "mm" 
       dpi = 200)
```

### Boxplot of all scores signatures{.tabset}

```{r Boxplot for each signature PTB vs. Control,results = "asis", message = FALSE, fig.height = 16, fig.width=24}
for (i in TBsig_PTB_Control){

  cat("####", i, "\n")
  
  get_boxplot(ssgsea_PTB_Control, sig_name = i)
  #print(signatureBoxplot(ssgsea_result, name = i, signatureColNames = i,
  #               annotationColName = c("Disease")))
  
  cat("\n\n")
}
```

#### Lee_4

```{r,results="hide",fig.height = 16, fig.width=24}
get_boxplot(ssgsea_PTB_Control[-index_Lee], sig_name = "Lee_4")
```

#### Rajan_HIV_5

```{r,results="hide",fig.height = 16, fig.width=24}
get_boxplot(ssgsea_PTB_Control[-index_Lee], sig_name = "Rajan_HIV_5")
```

#### Sloot_HIV_2

```{r,results="hide",fig.height = 16, fig.width=24}
get_boxplot(ssgsea_PTB_Control[-index_Lee], sig_name = "Sloot_HIV_2")
```

#### Thompson_FAIL_13

```{r,results="hide",fig.height = 16, fig.width=24}
get_boxplot(ssgsea_PTB_Control[-index_Lee], sig_name = "Thompson_FAIL_13")
```

#### Suliman_RISK_4

```{r,results="hide",fig.height = 16, fig.width=24}
get_boxplot(ssgsea_PTB_Control[-index_Suliman], sig_name = "Suliman_RISK_4")
```

## Positive TB (PTB) vs. Other Disease (OD){.tabset}
```{r calculate ssgsea for PTB vs OD, message = FALSE, results = "hide"}
multi_set_PTB_OD <- lapply(multi_data_all, function(x) get_sobject_TBSig(x,"PTB","OD")) %>% plyr::compact()

ssgsea_PTB_OD <- lapply(multi_set_PTB_OD,
                                 function(x) runTBsigProfiler(input = x,
                                                              useAssay = "counts",
                                                              signatures = TBsignatures,
                                                              algorithm = "ssGSEA",
                                                              combineSigAndAlgorithm = TRUE,
                                                              parallel.sz = 10))

```

### Boxplot of all scores signatures{.tabset}

```{r Boxplot for each signature PTB vs. OD,results = "asis", message = FALSE, fig.height = 16, fig.width=24}
for (i in names(TBsignatures)){

  cat("####", i, "\n")
  
  get_boxplot(ssgsea_PTB_OD, sig_name = i)
  #print(signatureBoxplot(ssgsea_result, name = i, signatureColNames = i,
  #               annotationColName = c("Disease")))
  
  cat("\n\n")
}
```

# Comparing Signature Scores via AUCs {.tabset}

## Positive TB (PTB) vs. Latent TB (Latent)

```{r AUC ridge plot for PTB vs Latent,fig.height = 15, fig.width=8, message=FALSE}
combine_auc <- function(result_list){
    aucs_result <- lapply(result_list, function(x)
    get_pvalue_auc(x,
                   annotationColName = "Disease",
                   signatureColNames = names(colData(x))[-c(1:2)]))
  aucs_result_dat <- do.call(rbind,aucs_result)
  
  # re-order data based on their median AUC
  aucs_result_dat_median <- aucs_result_dat %>% group_by(Signature) %>% summarise_all(median) %>% arrange(desc(AUC))
  
  ##########################
  
  Signature_order <- as.character(aucs_result_dat_median$Signature)
  # Re-order gene siganture, re-level
  aucs_result_dat$Signature <- factor(aucs_result_dat$Signature, levels = Signature_order)
  

  return(aucs_result_dat)
}

# Ridge plot for AUCs distribution across datasets
# PTB vs. Latent

get_auc_distribution(combine_auc(ssgsea_PTB_Latent)) + ggtitle("Ridge plot of AUC for PTB vs. Latent")

# p = get_auc_distribution(combine_auc(ssgsea_PTB_Latent)) + ggtitle("Ridge plot of AUC for PTB vs. Latent")
# ggsave("~/Desktop/auc.pdf",p,width = 8, height = 15, 
#       units = "in", # other options are "in", "cm", "mm" 
#       dpi = 200)
```

## Positive TB (PTB) vs. Health Control (Control)

```{r AUC ridge plot for PTB vs Control,fig.height = 15, fig.width=8}
# Ridge plot for AUCs distribution across datasets
# PTB vs. Health Control
get_auc_distribution(combine_auc(ssgsea_PTB_Control)) + ggtitle("Ridge plot of AUC for PTB vs. Control")

```

## Positive TB (PTB) vs. Other Disease (OD)

```{r AUC ridge plot for PTB vs OD,fig.height = 15, fig.width=8}
# Ridge plot for AUCs distribution across datasets
# PTB vs. OD

get_auc_distribution(combine_auc(ssgsea_PTB_OD)) + ggtitle("Ridge plot of AUC for PTB vs. OD")

```
